Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-24 paper/paper.tex
updating table

\Mar{----------------------------------- PAREI AQUI}

The goal of the first step is to extract options from the positive
traffic, which is characterized with pcap files~\cite{pcap}, a popular
format to describe network packets.

For that, \tname{} parses the message looking for options associated
with the data. As described in Section~\ref{sec:rules-and-packets},
there is a correspondence between packet fields and rule
options. \tname{} leverages that mapping to infer options. The
examples of sections~\ref{sec:dos} and~\ref{sec:active-recon} show
cases where \tname{} infers options with this mapping. Unfortunately,
inferring options at the field granularity level is insufficient to
capture content options as those described on
Section~\ref{sec:content-example}. These options refer to parts of the
payload of the message and cannot be extracted directly from the
fields. Handling this case is critically important as 93\% of the
rules contain \CodeIn{content} options.

Figure~\ref{fig:distribution-contents} shows the histogram of the
number of these options per rule obtained from a public
ruleset~\cite{emerging-threats-open} containing 27.8K Suricata
rules. We limited the number of content options per rule to 10 for
space. Rules without \CodeIn{content} options is an exception
(7\%) and half of the rules contain at least two
\CodeIn{content} options.  To find the values associated with these
options, \tname{} splits the payload of the malicious message in
tokens using natural language delimiters, such as $\backslash$t,
$\backslash$n, $\backslash$r, and spaces. We observed, by inspecting
existing rules, that \CodeIn{content} options typically refer to
sequence of characters in the payload of messages separated by these
delimiters. This tokenization step can generate dozens of options.
\tname{} produced a total of \Fix{xx} content options for the example
Section~\ref{sec:content-example}.

\Mar{Revisar.........................}
\Fix{
A avaliação que fazemos aqui pontua um dado content de acordo com
quantas vezes o campo em que ele se encontra eh referenciado nas
regras. Por exemplo, vamos supor que o tokenizer produziu a opção
content: "/HtmlAdaptor". Essa string pertence ao campo "Request
URI:". Esse campo pode ser acessado pelo modificador http\_uri. O
tokenizer associa o content a esse modificador.
}

@@

@@

\begin{table}[h!]
  \small
  \caption{\label{table:results}Results.}
  \centering
  \begin{tabular}{lcrp{9cm}}
    \toprule
    \multicolumn{1}{c}{Name} &
    \multicolumn{1}{c}{\#~Options Seed Rule} &
    \multicolumn{1}{c}{Rank} \\
    \midrule
    Ping Scan & 5 & - \\
    SYN Flood & 2 & - \\
    JBoss JMX & 23 & 23 \\
    WeBid & 23 & -\\    
    ColdFusion & 19 & 5\\
    .htaccess access & 17 & 2\\
    .idq access & 19 & 2 \\
    iisadmin access & 17 & 1 \\
    /system32/ in URI & 20 & 2 \\
    Script tag in URI & 22 & - \\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

The first step of \tname\ is Reverse Engineering. The goal of this
step is to extract options from the malicious traffic. A rule that
contains these options, by construction, captures the malicious
traffic. \tname\ uses two strategies to extract options: 1)~field and
2)~payload. We elaborate these strategies in the following.

For the field strategy, \tname{} parses the input messages looking for
options associated with the data in the fields of a network
message. Section~\ref{sec:rules-and-packets} shows the correspondence
between packet fields and some of the rule options from
\suri. Section~\ref{sec:active-recon} shows a concrete example where
the values for options are directly extracted from packet fields. We
constructed the mapping of network fields to options manually, from
the \suri\ documentation.

Unfortunately, the relevant data for the option may be at the message
payload not the fields. Figure~\ref{fig:rule-jboss} shows an example
of a rule that captures traffic containing the string
\CodeIn{"/HtmlAdaptor"} in the payload.  The content strategy focuses
on extracting (content) options from the payload. It is very important
to note that content options are prevalent. For example,
\percRulesWithContent{} of the \numrulessuri{} rules from the public
data set~\cite{emerging-threats-open} we analyzed contain at least one
option of this kind. Figure~\ref{fig:distribution-contents} shows the
histogram of the number of these options per rule for the data set
mentioned above. We limited the number of content options per rule to
10 for space. Rules without \CodeIn{content} options are an exception
and half of the rules contain at least two \CodeIn{content} options.
To find the values associated with these options, \tname{} splits the
payload of the malicious message in tokens using natural language
delimiters, such as $\backslash$t, $\backslash$n, $\backslash$r, and
spaces. We observed, by inspecting existing rules, that
\CodeIn{content} options typically refer to sequence of characters in
the payload of messages separated by these delimiters. This
tokenization step can generate dozens of options.  For example,
considering the example from Section~\ref{sec:content-example},
\tname{} produces a total of \Fix{XXX} content options.

\Mar{Nao esta muito claro como isto funciona. Rodamos sempre as duas
  estrategias (para obter field options e content options)?
  (Precisamos dizer se eh possivel ter regras com os dois tipos de
  opcoes). Como isto funciona 1) quando uma ataque tem varias
  mensagens (e.g., ping scan) e 2) quando temos varios ataques?}

\textbf{Implementation.}\Fix{elaborate...}
, characterized with pcap files~\cite{pcap}, a popular format
to describe network packets.

%% \Fix{
%% A avaliação que fazemos aqui pontua um dado content de acordo com
%% quantas vezes o campo em que ele se encontra eh referenciado nas
%% regras. Por exemplo, vamos supor que o tokenizer produziu a opção
%% content: "/HtmlAdaptor". Essa string pertence ao campo "Request
%% URI:". Esse campo pode ser acessado pelo modificador http\_uri. O
%% tokenizer associa o content a esse modificador.
%% }

@@

\Mar{------------------------------------------- Eu estou aqui}

@@

\begin{table}[h!]
  \small
  \caption{\label{table:results}Results.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrr}
    \toprule
    \multicolumn{1}{c}{\multirow{2}{*}{Name}} &
    \#~Options &
    \#~Plausible Rules &    
    \multicolumn{1}{c}{Rank} \\

     &
    \multicolumn{1}{c}{(step 1)} &
    \multicolumn{1}{c}{(step 2)} &    
    \multicolumn{1}{c}{(step 3)} \\

    \midrule
    Ping Scan & 5 & - & - \\
    SYN Flood & 2 & - & - \\
    JBoss JMX & 23 & - & 23 \\
    WeBid & 23 & - & -\\    
    ColdFusion & 19 & - & 5\\
    .htaccess access & 17 & - & 2\\
    .idq access & 19 & - & 2 \\
    iisadmin access & 17 & - & 1 \\
    /system32/ in URI & 20 & - & 2 \\
    Script tag in URI & 22 & - & - \\
    \bottomrule
  \end{tabular}
\end{table}
